Super vs fficient absorption of light by spherical nanoparticles

### 2015-07-06 2015-Ladutenko-Qabs/2015-Ladutenko-Qabs-aem.tex
Changes from Andrey

\title{Super absorption and efficient absorption of light by spherical
  nanoparticles}

@@

\begin{abstract}
  Spherical subwavelength nanoparticles have a fundamental limit as to
  how much light they can absorb. This limit is based on the
  assumption that only one mode is excited in the nanoparticle. Using
  genetic algorithm, we design multi-layer nanoparticles, in which we
  can make several resonant modes overlap at the same frequency, thus
  significantly beating the theoretical limit of absorption, and we
  call this super-absorption. We further introduce the efficiency of
  the absorption for a nanoparticle, which is absorption normalized by
  the physical size of the particle, and show that efficient absorbers
  are not always operating in super-absorbing regime.
\end{abstract}

@@

Mie theory~\cite{Mie-1908}, which is over 100 years old now, describes
interaction of electromagnetic waves with spherical particles. Mie
solution is still of great interest these
days~\cite{Suzuki-2008,MacKowski-2012,Lerme-2000,Xu-2005,Li-2006,Gogoi-2010,Santiago-2011},
since it is one of the primary tools for analyzing wave scattering by
spherical objects. Further development of the Mie
theory~\cite{Yang-2003, Pena-scattnlay-2009} made it possible to apply
study the multilayered spherical
particles~\cite{Sheehan-2013,Selmke-2012}.  Such particles have
various applications in cancer treatment~\cite{Zhang-2010,
  Hirsch-2003} and medical diagnostics~\cite{Allain-2002},
cloaking~\cite{Qui-2009, Semouchkina-2013, Ladutenko-2014} and
plasmonic~\cite{Martin-2013, Alu-2005} devices, study on thermal
properties of insulating material~\cite{Xie-2013}, as well as for
improving solar cells performance~\cite{Kameya-2011,Mann-2011}, and so
on.

The scattering properties of multilayer cylinders and spheres was
studied in great detail by Fan~et~al.~\cite{Fan-2010,Fan-2011}.  In these
works authors introduced the concept of a super scattering, when the
scattering cross-section of a multilayer particle exceeds that of a
homogeneous particle of the same size in the so-called single-channel
limit. The super scattering appears when we design a multilayer
structure so that several modes become nearly degenerate, i.e. their
resonance frequencies coincide or get close to each other. In a
homogeneous particle, the resonances appear at different frequencies,
and there is no design freedom for a fixed geometry of the structure
to make these resonances overlap, and this limits the achievable
scattering cross-section.

@@

To overcome these limitations, we employ similar approach for
enhancing absorption cross-section~\cite{Fan-2011}. In particular, we
propose to use the multi-layer structures, and by means of genetic
algorithm we optimize the ACS of these particles. We analyze the
absorption cross-section of these particles, and present the
super-absorption regime. We further introduce the absorption
efficiency, which is the ACS normalized to the geometric cross-section
of the particles. Here we show that there is a strong counter-play
between the increased absorption for larger particles vs size for
smaller particles, and quite remarkably we find that the most {\em
  efficient} absorption can be achieved in a single channel limit for
small particles.

@@

It is obvious that in general, a bigger particle will have a bigger
absorption cross-section, so sphere with the diameter of 1 cm will
absorb more light than any sphere at the nanoscale. Therefore, we use
absorption efficiency $Q_{\rm abs}=C_{\rm abs}/2\pi R_{\rm total}^2$,
where $R_{\rm total}$ is the outer radius of the particle and $C_{\rm
  abs}$ is the absorption cross-section. We maximize absorption
efficiency at a fixed wavelength of incident light (we have chosen
$\lambda=500$~nm).

@@

      \label{fig:overview}
    }%
  }
\end{figure}
%
In order to study the dependence of the absorption efficiency on the
outer particle size, we run optimization algorithm for different
(fixed) particle outer size, and our optimization parameters are the
radii of internal cores, whereas the target function is the absorption
efficiency.  We show the results of our genetic optimization algorithm
in Fig.~\ref{fig:overview}a.  Dashed lines show theoretical absorption
limit of a dipole ($n=1$) and a quadrupole ($n=2$)
resonances~\cite{Tribelsky-2011}, which are given as $$Q^{(n)}_{\rm abs\
  max}=\frac{2n+1}{2q^2},$$ where the size parameter $q=2\pi R_{\rm
  total}/\lambda$, and $n$ is an angular momentum of the
mode. Following Ref.~\cite{Fan-2011}, where authors introduce super
scattering for spherical particles, here we we introduce
super-absorption, when the ACS is larger than the theoretical limit
for absorption by the mode with highest excited angular momentum
$n$. In our parameter space we have just modes up to the quadrupole
excited ($n=2$), and in order to get a super-absorption our efficiency
should be higher than that of a quadrupole. We clearly see this
super-absorption at $R_{\rm total}>60$~nm, in Fig.~\ref{fig:overview}
(a).

@@

Fig.~\ref{fig:overview} (c) shows optimized sizes of the layers inside
the multi-layer structure. It reveals quite a curious result, that the
dipole branch (i.e. for particle radii below 56.6~nm) has two
parts. For $R_{\rm total}<46$~nm the best absorber has just two
layers, as the radius of the core of the three-layer structure
vanishes, and the particle reduces to $Ag/Si$ core-shell structure.
At $R_{\rm total}=46$ dipole channel becomes practically
undistinguishable from the theoretical limit (it becomes
$\tilde{a}_1>0.249$).  It appears that the optimizer introduced the
inner $Si$ layer in order to keep $\tilde{a}_1$ near the theoretical
limit as the $R_{\rm total}$ increases.  As a side effect, the
quadrupole contribution $\tilde{a}_2$ appears, however, it does not
help to reach super-absorption limit $n=2$.

@@

\begin{figure}
  \center{
    % Use pdfcrop to remove white margins
    \includegraphics[width=0.4\textwidth]{fig/2015-04-01-SiAgSi-ab-spectra3.pdf}%
    \caption{Spectra of Mie absorption coefficients of (a) efficient and (b-c)super absorption design.      
      \label{fig:spectra}
    }%
  }
\end{figure}
%
To study spectral properties of the structures with large ACS which we
obtained by the optimization, in Fig.~\ref{fig:spectra} we plot three
different cases for designs that correspond to local maxima of $Q_{\rm
  abs}$ shown in Fig.~\ref{fig:overview} (a).  As expected, the design
corresponding to the maximum absorption efficiency at R=36~nm has a
single electric dipole resonance centered at the target wavelength
$\lambda=500$~nm. Spectra of designs with maxima at R=63~nm and
R=81~nm have a signature of the super absorption, i.e. there is an
overlap of several resonances.  We note that these structures have
additional absorption resonances, but they are located far from the
wavelength of interest.

@@

\begin{figure}
  \center{
    % Use pdfcrop to remove white margins
    \includegraphics[width=0.4\textwidth]{fig/SiAgSi-flow-R62-YZ-Eabs.pdf}%
    \caption{ Amplitude of electric field for R=36 nm (a,c) and
      ``super'' (b,d) designs in E-k  (a-b) and H-k (c-d) planes. 
      \label{fig:field}
    }%
  }
\end{figure}
%
Finally, we present distribution of the amplitude of the electric
field in Fig.~\ref{fig:field} for two designs: with the best
efficiency at $R=36$~nm and in a super-absorbing design with
$R=63$~nm.  We set different scale for these two designs so that black
circles that denote outer boundary are plotted to be the same size.
We also plot streamlines of the Poynting vector which characterize
energy flow. For the effective design of the absorber, the power from
a large cross-sectional area flows into the particle.  In case of
super-absorption regime, we observe power flow vortices, which make
absorbtion more efficient as the electromagnetic energy propagates
several times through the absorbing materials.

In conclusion, we find the effect of the super-absorption, when the
absorption cross-section of the nanoparticle can exceed the
theoretical limit for the absorption by the highest excited mode. This
occur when several resonance modes of the structure overlap at the
same frequency. We introduce efficiency of the absorption as an
absorption cross-section divided by a geometric cross-section of the
particle, and quite unexpectedly we find that the most efficient
absorbers are smaller nanoparticles working in a single mode
regime. We present their spectral characteristics and field structure.

>>>>

\title{Super vs fficient absorption of light by spherical nanoparticles}

@@

\begin{abstract}
  Nanoparticles have a fundamental limit as to
  how much light they can absorb. This limit is based on the
  finite number of modes excited inside the nanoparticle at a given wavelength and maximal absorption capacity per mode. Thus, the enhanced absorption can be achieved when each mode supported by the nanopartice absorbs light up to the maximal capacity. Using genetic algorithm, we design multi-layer nanoparticles, in which we
  can make several resonant modes overlap at the same frequency resulting in {\it super-absorption}. 
  We further introduce the {\it efficiency of
  the absorption} for a nanoparticle, which is absorption normalized by
  the physical size of the particle, and show that efficient absorbers
  are not always operating in super-absorbing regime.
\end{abstract}

@@

Mie theory~\cite{Mie-1908}, which is over 100 years old now, describes
interaction of electromagnetic waves with spherical particles. Mie
solution is still of great current interest~\cite{Suzuki-2008,MacKowski-2012,Lerme-2000,Xu-2005,Li-2006,Gogoi-2010,Santiago-2011},
since it is one of the primary tools for analyzing wave scattering by
spherical objects. Further development of the Mie
theory~\cite{Yang-2003, Pena-scattnlay-2009} made it possible to apply
study the multilayered spherical
particles~\cite{Sheehan-2013,Selmke-2012}.  Such particles have
various applications in cancer treatment~\cite{Zhang-2010,
  Hirsch-2003} and medical diagnostics~\cite{Allain-2002},
cloaking~\cite{Qui-2009, Semouchkina-2013, Ladutenko-2014} and
plasmonic~\cite{Martin-2013, Alu-2005} devices, study on thermal
properties of insulating material~\cite{Xie-2013}, as well as for
improving solar cells performance~\cite{Kameya-2011,Mann-2011}, and so
on.

The scattering properties of multilayer cylinders and spheres was
studied in great detail by Fan~et~al.~\cite{Fan-2010,Fan-2011}.  In these
works authors introduced the concept of a super-scattering, when the
scattering cross-section of a multilayer particle exceeds that of a
homogeneous particle of the same size in the so-called single-channel
limit. The super scattering appears when we design a multilayer
structure so that several modes become nearly degenerate, i.e. their
resonance frequencies coincide or get close to each other. In a
homogeneous particle, the resonances appear at different frequencies,
and there is no design freedom for a fixed geometry of the structure
to make these resonances overlap, and this limits the achievable
scattering cross-section.

@@

To overcome these limitations, we employ similar approach for
enhancing absorption cross-section~\cite{Fan-2011}. In particular, we
propose to use the multi-layer structures, and by means of genetic
algorithm we optimize the ACS of such nanoparticles. We analyze the
absorption cross-section of these particles, and present the
super-absorption regime. We further introduce the absorption
efficiency, which is the ACS normalized to the geometric cross-section
of the particles. Here we show that there is a strong counter-play
between the increased absorption for larger particles vs size for
smaller particles, and quite remarkably we find that the most {\em
  efficient} absorption can be achieved in a single channel limit for
small particles.

@@

It is a common understanding, that, in general, a bigger particle will have a bigger
absorption cross-section, so sphere with the diameter of 1 cm will
absorb more light than any nanoscale sphere at a give wavelength. Therefore, it is practical  to employ
the absorption efficiency $Q_{\rm abs}=C_{\rm abs}/2\pi R_{\rm total}^2$,
where $R_{\rm total}$ is the outer radius of the particle and $C_{\rm
  abs}$ is the absorption cross-section. We maximize absorption
efficiency at a fixed wavelength of the incident light (we have chosen
$\lambda=500$~nm).

@@

      \label{fig:overview}
    }%
  }
\end{figure}
%
In order to study the dependence of the absorption efficiency on the
outer particle size, we run optimization algorithm for different
(fixed) particle outer size, and our optimization parameters are the
radii of internal cores, whereas the target function is the absorption
efficiency.  We show the results of our genetic optimization algorithm
in Fig.~\ref{fig:overview}a.  Dashed lines show theoretical absorption
limit of a dipole ($n=1$) and a quadrupole ($n=2$)
resonances~\cite{Tribelsky-2011}, which are given as $$Q^{(n)}_{\rm abs\
  max}=\frac{2n+1}{2q^2},$$ where the size parameter $q=2\pi R_{\rm
  total}/\lambda$, and $n$ is an angular momentum of the
mode. Following Ref.~\cite{Fan-2011}, where authors introduced a super
scattering for spherical particles, here we introduce
super-absorption. In our parameter space we have just modes up to the quadrupole
excited ($n=2$), and in order to get a super-absorption our efficiency
should be higher than that of a quadrupole. We clearly see this
super-absorption at $R_{\rm total}>60$~nm, in Fig.~\ref{fig:overview}
(a).

@@

Fig.~\ref{fig:overview} (c) shows optimized sizes of the layers inside
the multi-layer structure. It reveals quite a curious result, that the
dipole branch (i.e. for particle radii below 56.6~nm) has two
parts. For $R_{\rm total}<46$~nm the best absorber has just two
layers, as the radius of the core of the three-layer structure
vanishes, and the particle reduces to $Ag/Si$ core-shell structure.
At $R_{\rm total}=46$ dipole channel reaches its theoretical limit (it becomes
$\tilde{a}_1>0.249$).  It appears that the optimizer introduced the
inner $Si$ layer in order to keep $\tilde{a}_1$ near the theoretical
limit as the $R_{\rm total}$ increases.  As a side effect, the
quadrupole contribution $\tilde{a}_2$ appears, however, it does not
help to reach super-absorption limit $n=2$.

@@

\begin{figure}
  \center{
    % Use pdfcrop to remove white margins
    \includegraphics[width=0.4\textwidth]{fig/2015-04-01-SiAgSi-ab-spectra3.pdf}%
    \caption{Spectra of Mie absorption coefficients of (a) efficient and (b-c)super absorption design.      
      \label{fig:spectra}
    }%
  }
\end{figure}
%
To study spectral properties of the structures with large ACS, which we
obtained by the optimization, in Fig.~\ref{fig:spectra} we plot three
different cases for designs that correspond to local maxima of $Q_{\rm
  abs}$ shown in Fig.~\ref{fig:overview} (a).  As expected, the design
corresponding to the maximum absorption efficiency at R=36~nm has a
single electric dipole resonance centered at the target wavelength
$\lambda=500$~nm. Spectra of designs with maxima at R=63~nm and
R=81~nm have a signature of the super absorption, i.e. there is an
overlap of several resonances.  We note that these structures have
additional absorption resonances, but they are located far from the
wavelength of interest.

@@

\begin{figure}
  \center{
    % Use pdfcrop to remove white margins
    \includegraphics[width=0.4\textwidth]{fig/SiAgSi-flow-R62-YZ-Eabs.pdf}%
    \caption{ Amplitude of electric field for R=36 nm (a,c) and
      ``super'' (b,d) designs in E-k  (a-b) and H-k (c-d) planes. 
      \label{fig:field}
    }%
  }
\end{figure}
%
Finally, we present distribution of the amplitude of the electric
field in Fig.~\ref{fig:field} for two designs: with the best
efficiency at $R=36$~nm and in a super-absorbing design with
$R=63$~nm.  We set different scale for these two designs so that black
circles that denote outer boundary are plotted to be the same size.
We also plot streamlines of the Poynting vector which characterize
energy flow. For the effective design of the absorber, the power from
a large cross-sectional area flows into the particle.  In case of
super-absorption regime, we observe power flow vortices, which make
absorption more efficient as the electromagnetic energy propagates
several times through the absorbing materials.

In conclusion, we find the effect of the super-absorption, when the
absorption cross-section of the nanoparticle can reach the
theoretical limit for several modes at once, corresponding to  the super-absorption condition. Such situation realized when  several resonant modes of the structure overlap at the
same frequency. Moreover, quite unexpectedly we find that the most efficient
absorbers, which are characterized by enhanced absorption efficiencies, are smaller nanoparticles working in a single mode
regime. We present their spectral characteristics and field structure.
